Equalizer Programming Guide

### 2008-05-16 Developer/ProgrammingGuide/paper.tex
Fill in more file format content in the Programming Guide.
[ ] May break build
[ ] Breaks existing applications
[ ] Bugfix
[ ] New Feature
[ ] Cleanup
[ ] Optimization
[x] Documentation

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-13, 2008 & File Format Appendix [incomplete]\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\subsubsection{Compound Channels}
Each compound has a channel, which is used by the compound to execute
the rendering tasks. One channel might be used by multiple
compounds. Unused channels are not instantiated during
initialization. The rendering tasks for the channels are computed by the
server and send to the appropriate render clients.

@@

\subsubsection{Compound Classification}
The channels of the leaf compounds in the compound tree are designated
as source channels. The top-most channel in the tree is the destination
channel. Only source channels execute rendering tasks. All channels in a
compound tree work for the destination channel. The destination channel
defines the 2D pixel viewport rendered by all leaf compounds. The
destination channel and pixel viewport can not be overridden by child
compounds.

@@

\subsubsection{Recomposition - Frames}
Compounds use output and input frames to configure the recomposition of
the resulting pixel data from the source channels. An output frame
connects to an input frame of the same name. The selected frame buffer
data is transported from the output channel to the input channel. The
assembly routine of the input channel will block on the availability of
the output frame. This composition process is extensively described in
\sref{sCompositing}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \subfigure[]{\label{fSync}
    \includegraphics[width=.25\textwidth]{images/sync}}\hfil
  \subfigure[]{\label{fAsync}
    \includegraphics[width=.25\textwidth]{images/async}}%
  {\caption{\small\label{fSyncAsync}Synchronous and asynchronous execution}}
\end{wrapfigure}
When the configuration was successfully initialized, the main rendering
loop is executed. It runs until the user exits the
configuration, or when a maximum number of frames has been rendered,
specified by a command-line argument. The latter is useful for
benchmarks. The \textsf{Clock} is reused for measuring the overall
performance. A new frame is started using \textsf{Config::startFrame}
and a frame is finished using \textsf{Config::finishFrame}.

@@

By default, \textsf{Config::finish\-Fra\-me} also synchronizes the
completion of all local rendering tasks for the current frame. This
facilitates porting of exising rendering codes, since the database does
not have to be multi-buffered. Applications such as \textsf{eqPly}, which
do not need this per-node frame synchronization can disable it, as
explained in \sref{sNodeFrame}.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/modelDist.pdf}
  {\caption{\small\label{fModelDist}Scene Data in eqPly}}
\end{wrapfigure}
\textsf{eqPly} uses static distributed objects to distribute the model
loaded by the application. This model can be easily extended to
versioned objects, in order to support dynamic data changes.

@@

The identifer of the root object of this distributed hierarchy is passed
as part of the \textsf{InitData}, and used during remote node
initialization to map the model. \fig{fModelDist} shows the UML
hierarchy of the model and distribution classes.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/objects.pdf}
  {\caption{\small\label{fObjects}Distributed Data Handling}}
\end{wrapfigure}
Applications distributing a dynamic scene graph use the frame data
instead of the init data as the entry point to their scene graph data
structure. \fig{fObjects} shows one possible implementation, where the
identifier and version of the scene graph root are transported using the
frame data. The scene graph root then serializes and de-serializes his
immediate children by transferring their identifier and current version,
similar to the static distribution done by \textsf{eqPly}.

The objects are still created by the application, and then registered or
mapped with the session in order to distribute them. When mapping
objects in a hierarchical data structure, their type often has to be
known in order to create them. Equalizer does not currently provide
object typing, this has to be done by the application, either implicitly
in the current implementation context, or by transferring a type
identifier. In \textsf{eqPly}, object typing is implicit since it is
well-defined which object is mapped in which context.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/windowInit.pdf}
  {\caption{\small\label{fWindowInit}Window Initialization}}
\end{wrapfigure}
\fig{fWindowInit} shows a flow chart of the window initialization. The
colored functions are task methods and can be replaced by
application-specific implementations.

@@

The state object is used to handle the creation of OpenGL objects in a
multipipe, multi-threaded execution environment. It uses an object
manager, which is described in detail in \sref{sObjectManager}. It is
used in conjunction with a reference-counting smart-pointer here, since it is
potentially `owned' by multiple windows at the same time.

@@

After the basic view setup, a directional light is configured, and the
model is positioned using the camera parameters from the frame data. The
camera parameters are transported using the the frame data to ensure
that all channels render a given frame using the same position.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/DB.png}
  {\caption{\small\label{fDB}Destination view of a DB compound}}
\end{wrapfigure}
In order to draw the model, a helper class for view frustum culling is
set up using the view frustum from Equalizer and the camera position
from the frame data. The frustum helper computes the six frustum planes
from the projection and modelView matrices. During rendering, the
bounding spheres of the model are tested against these planes to
determine the visibility with the frustum.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\small\label{fRender}Main Rendering Loop}}
\end{wrapfigure}
The model data is spatially organized in an 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree and each
node of the tree gets a database range assigned. The root node
has the range [0, 1], its left child [0, 0.5] and its right child [0.5,
1], and so on for all nodes in the tree. The preprocessed model is saved
in a binary format for accelerating subsequent use.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\small\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
Please note that the default thread synchronization synchronizes all
\textsf{Channel::frame\-Draw} operations on a \textbf{single} node with
the node's main thread. The per-node frame synchronization does not
break the asynchronous execution across nodes. Advanced applications can
remove the per-node frame synchronization.

@@

Each entity uses \textsf{waitFrameStarted} to block on the parent's
\textsf{startFrame}, e.g., \textsf{Pipe::\-frame\-Start} calls
\textsf{Node::wait\-Frame\-Started} to wait for the corresponding
\textsf{Node::start\-Fra\-me}. This explicit synchronization allows to
update non-critical data before synchronizing with
\textsf{waitFrameStarted}, or after unlocking using
\textsf{start\-Fra\-me}. \fig{fFrameSync} illustrates this
synchronization model.

@@

DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
Finding the correct assembly order for perspective frustra is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frustra. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

@@

Statistics are measured in milliseconds since the configuration was
initialized. On each node, the global configuration clock is reset at
the same time by the server during \textsf{Config::init}. Each statistic
event records the originator's (channel, window or config) unique
identifier.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/statistics.png}
  {\caption{\small\label{fStatistics}Statistics for a two-node 2D compound}}
\end{wrapfigure}
Statistics are enabled per entity using an attribute hint. The hint
determines how precise the gathered statistics are. When set to
\textsf{fastest}, the per-frame clock is sampled directly when the event
occurs. When set to \textsf{nicest}, all OpenGL commands will be
finished before sampling the event. This incurs a performance penalty,
but gives more correct results. The default setting is fastest in
release builds, and nicest in debug builds.

@@

\subsection{Global Section}

@@

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DRAWABLE\\
\textbf{Value} & window \textbar \ pbuffer\\
\textbf{Default} & window\\
\textbf{Details} & Selects the window's drawable type.\\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
\textbf{Value} & off \textbar \ fastest [on] \textbar \ nicest\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_COLOR\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_ALPHA\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_DEPTH\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_STENCIL\\
\textbf{Value} & unsigned\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
\textbf{Value} & OFF \textbar \ FASTEST [ ON ] \textbar \ NICEST\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
\textbf{Value} & QUAD \textbar \ ANAGLYPH\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_LEFT\_MASK\\
\textbf{Value} & [ RED GREEN BLUE ]\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_RIGHT\_MASK\\
\textbf{Value} & [ RED GREEN BLUE ]\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
\hline
\textbf{Name} & EQ\_COMPOUND\_IATTR\_UPDATE\_FOV\\
\textbf{Value} & HORIZONTAL [ON] \textbar \ VERTICAL \textbar \ OFF\\
\textbf{Default} & \\
\textbf{Details} & \\
\hline
\end{tabularx}\\
\end{center}

@@

{\footnotesize\begin{lstlisting}
server
{
    connection // 0-n times, listening connections of the server
    {
        type       TCPIP | SDP
        TCPIP_port unsigned
        hostname   string
    }

@@

For an introduction on the concepts of nodes, pipes, windows, channels
and compounds please refer to \sref{sConfig} and \sref{sCompounds}.

@@

@@

{\footnotesize\begin{lstlisting}
{
    name     string
    connection // 0-n times, possible connections to this node
    {
        type          TCPIP | SDP
        TCPIP_port    unsigned
        hostname      string
        command       string      // render client launch command
        command_quote 'character' // launch command argument quote char
        timeout       unsigned    // timeout in milliseconds for launch
    }
    pipe {...} // 1-n times
}
\end{lstlisting}}

>>>>

\clearpage
\thispagestyle{empty}
\tableofcontents
\vfill{\center\begin{tabularx}{\textwidth}{|l|l|X|}
    \hline
    \bf Rev & \bf Date     & \bf Changes \\
    \hline
    1.0     & Oct 28, 2007 & Initial Version for Equalizer 0.4\\
    1.2     & Apr 15, 2008 & Version for Equalizer 0.5\\
    1.3.1   & May 02, 2008 & Add statistics, improved images based on 1.2 print\\
    1.3.2   & May 05, 2008 & Add network distribution on eqPly model data\\
    1.3.3   & May 10-16, 2008 & File Format Appendix [incomplete]\\
    \hline
  \end{tabularx}}
\thispagestyle{empty}
\clearpage

@@

\subsubsection{Compound Channels}
Each compound has a channel, which is used by the compound to execute
the rendering tasks. One channel might be used by multiple
compounds. Unused channels, windows, pipes and nodes are not
instantiated during config initialization. The rendering tasks for the
channels are computed by the server and send to the appropriate render
client nodes at the beginning of each frame.

@@

\subsubsection{Compound Classification}
The channels of the leaf compounds in the compound tree are designated
as source channels. The top-most channel in the tree is the destination
channel. Only source channels execute the draw task. All channels in a
compound tree work for the destination channel. The destination channel
defines the 2D pixel viewport rendered by all leaf compounds. The
destination channel and pixel viewport can not be overridden by child
compounds.

@@

\subsubsection{Recomposition - Frames}
Compounds use output and input frames to configure the recomposition of
the resulting pixel data from the source channels. An output frame
connects to an input frame of the same name. The selected frame buffer
data is transported from the output channel to the input channel. The
assembly routine of the input channel will block on the availability of
the output frame. This composition process is extensively described in
\sref{sCompositing}. Frame names are only valid within the compound
tree, that is, an output frame from one compound tree can not be used as
an input frame of another compound tree.

\subsubsection{Synchronization - Swapbarrier}
Compounds may have a swapbarrier to synchronize the buffer swap of
multiple channels. Windows with a swapbarrier of the same name enter a
barrier before executing the swap buffer task. Before entering the
barrier, \textsf{Window::finish} is called to ensure that all OpenGL
commands have been executed.

Note that this software swap barrier is not accurate enough for
edge-blended projection system. Such installations typically use a
hardware synchronization, e.g., nVidia G-Sync cards. It is however
sufficient for display walls made out of LCD's.

Swapbarrier names are only valid within the compound tree, that is, a
compound from one compound tree can not be synchronized with a compound
from another compound tree.

@@

When the configuration was successfully initialized, the main rendering
loop is executed. It runs until the user exits the
configuration, or when a maximum number of frames has been rendered,
specified by a command-line argument. The latter is useful for
benchmarks. The \textsf{Clock} is reused for measuring the overall
performance. A new frame is started using \textsf{Config::startFrame}
and a frame is finished using \textsf{Config::finishFrame}.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \subfigure[]{\label{fSync}
    \includegraphics[width=.25\textwidth]{images/sync}}\hfil
  \subfigure[]{\label{fAsync}
    \includegraphics[width=.25\textwidth]{images/async}}%
  {\caption{\small\label{fSyncAsync}Synchronous and asynchronous execution}}
\end{wrapfigure}
By default, \textsf{Config::finish\-Fra\-me} also synchronizes the
completion of all local rendering tasks for the current frame. This
facilitates porting of exising rendering codes, since the database does
not have to be multi-buffered. Applications such as \textsf{eqPly}, which
do not need this per-node frame synchronization can disable it, as
explained in \sref{sNodeFrame}.

@@

\textsf{eqPly} uses static distributed objects to distribute the model
loaded by the application. This model can be easily extended to
versioned objects, in order to support dynamic data changes.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/modelDist.pdf}
  {\caption{\small\label{fModelDist}Scene Data in eqPly}}
\end{wrapfigure}
The identifer of the root object of this distributed hierarchy is passed
as part of the \textsf{InitData}, and used during remote node
initialization to map the model. \fig{fModelDist} shows the UML
hierarchy of the model and distribution classes.

@@

Applications distributing a dynamic scene graph use the frame data
instead of the init data as the entry point to their scene graph data
structure. \fig{fObjects} shows one possible implementation, where the
identifier and version of the scene graph root are transported using the
frame data. The scene graph root then serializes and de-serializes his
immediate children by transferring their identifier and current version,
similar to the static distribution done by \textsf{eqPly}.

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/objects.pdf}
  {\caption{\small\label{fObjects}Scene Graph Distribution}}
  \vspace{-4em}
\end{wrapfigure}
The objects are still created by the application, and then registered or
mapped with the session in order to distribute them. When mapping
objects in a hierarchical data structure, their type often has to be
known in order to create them. Equalizer does not currently provide
object typing, this has to be done by the application, either implicitly
in the current implementation context, or by transferring a type
identifier. In \textsf{eqPly}, object typing is implicit since it is
well-defined which object is mapped in which context.

@@

\fig{fWindowInit} shows a flow chart of the window initialization. The
colored functions are task methods and can be replaced by
application-specific implementations.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/windowInit.pdf}
  {\caption{\small\label{fWindowInit}Window Initialization}}
%  \vspace{-1em}
\end{wrapfigure}
The state object is used to handle the creation of OpenGL objects in a
multipipe, multi-threaded execution environment. It uses an object
manager, which is described in detail in \sref{sObjectManager}. It is
used in conjunction with a reference-counting smart-pointer here, since it is
potentially `owned' by multiple windows at the same time.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/DB.png}
  {\caption{\small\label{fDB}Destination view of a DB compound}}
\end{wrapfigure}
After the basic view setup, a directional light is configured, and the
model is positioned using the camera parameters from the frame data. The
camera parameters are transported using the the frame data to ensure
that all channels render a given frame using the same position.

@@

In order to draw the model, a helper class for view frustum culling is
set up using the view frustum from Equalizer and the camera position
from the frame data. The frustum helper computes the six frustum planes
from the projection and modelView matrices. During rendering, the
bounding spheres of the model are tested against these planes to
determine the visibility with the frustum.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \vspace{-1em}
  \includegraphics[width=.618\textwidth]{images/render.pdf}
  {\caption{\small\label{fRender}Main Render Loop}}
  \vspace{-2em}
\end{wrapfigure}
The model data is spatially organized in an 3-dimensional
kD-tree\footnote{See also \link{http://en.wikipedia.org/wiki/Kd-tree}}
for efficient view frustum culling. When the model is loaded by
\textsf{Node::config\-Init}, it is preprocessed into the kD-tree and each
node of the tree gets a database range assigned. The root node
has the range [0, 1], its left child [0, 0.5] and its right child [0.5,
1], and so on for all nodes in the tree. The preprocessed model is saved
in a binary format for accelerating subsequent use.

@@

Please note that the default thread synchronization synchronizes all
\textsf{Channel::frame\-Draw} operations on a \textbf{single} node with
the node's main thread. The per-node frame synchronization does not
break the asynchronous execution across nodes. Advanced applications can
remove the per-node frame synchronization.

@@

\begin{wrapfigure}{r}{.618\textwidth}
  \includegraphics[width=.618\textwidth]{images/frameSync.pdf}
  {\caption{\small\label{fFrameSync}Per-Node Frame Synchronization}}
\end{wrapfigure}
Each entity uses \textsf{waitFrame\-Started} to block on the parent's
\textsf{startFrame}, e.g., \textsf{Pipe::\-frame\-Start} calls
\textsf{Node::wait\-Frame\-Started} to wait for the corresponding
\textsf{Node::start\-Fra\-me}. This explicit synchronization allows to
update non-critical data before synchronizing with
\textsf{waitFrameStarted}, or after unlocking using
\textsf{start\-Fra\-me}. \fig{fFrameSync} illustrates this
synchronization model.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/volResult.png}
  {\caption{\small\label{fVolResult}Result of \fig{fBlendPerspective}}}
\end{wrapfigure}
DB compounds have the advantage of scaling any part of the volume
rendering pipeline: texture and main memory (smaller bricks for each
channel), fill rate (less samples per channel) and IO bandwidth for
time-dependent data (less data per time step and channel). Since the
amount of texture memory needed for each node decreases linearly, they
make it possible to render data sets which are not feasible to
visualize with any other approach.

@@

Finding the correct assembly order for perspective frustra is more
complex. The perspective distortion invalidates a simple orientation
criteria like the one used for orthographic frustra. For the view and
frustum setup shown in \fig{fBlendPerspective}\footnote{Volume Data Set
  courtesy of: AVS, USA} the correct compositing order is 4-3-1-2 or
1-4-3-2.

@@

\begin{wrapfigure}{r}{.382\textwidth}
  \includegraphics[width=.382\textwidth]{images/statistics.png}
  {\caption{\small\label{fStatistics}Statistics for a two-node 2D compound}}
\end{wrapfigure}
Statistics are measured in milliseconds since the configuration was
initialized. On each node, the global configuration clock is reset at
the same time by the server during \textsf{Config::init}. Each statistic
event records the originator's (channel, window or config) unique
identifier.

Statistics are enabled per entity using an attribute hint. The hint
determines how precise the gathered statistics are. When set to
\textsf{fastest}, the per-frame clock is sampled directly when the event
occurs. When set to \textsf{nicest}, all OpenGL commands will be
finished before sampling the event. This incurs a performance penalty,
but gives more correct results. The default setting is fastest in
release builds, and nicest in debug builds.

@@

\subsection{\label{sGlobal}Global Section}

@@

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_DRAWABLE\\
  \textbf{Value} & window \textbar \ pbuffer\\
  \textbf{Default} & window\\
  \textbf{Details} & Selects the window's drawable type. A window is an
  on-screen, window system-dependent window with a full-window OpenGL
  drawable. PBuffers are off-screen drawables created using window
  system-dependent PBuffer API's. In order to calculate the PBuffer size
  on unconnected devices, a pipe viewport size of 4096x4096 is assumed,
  unless specified otherwise.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ON] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & Determines how statistics are gathered. OpenGL
  buffers commands, which causes the rendering to be executed at an
  arbitrary point in time. Nicest statistics gathering executes a
  \textsf{Window::finish}, which calls by default \textsf{glFinish}, in
  order to accurately account the rendering operations to the sampled
  task method. However, calling \textsf{glFinish} has a performance
  impact. Therefore, the fastest statistics gathering samples the task
  statistics directly, without finishing the OpenGL commands.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_COLOR\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the number of color planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of color
  planes specified. AUTO select a visual with a reasonable quality,
  typically eight bits per color.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_ALPHA\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of alpha planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of alpha
  planes specified. By default no alpha planes are requested.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_DEPTH\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & AUTO\\
  \textbf{Details} & Determines the precision of the depth buffer. The
  interpretation of this value is window system-specific, as some window
  systems select a visual with the closest match to this value, and some
  select a visual with at least the number of depth bits specified. AUTO
  select a visual with a reasonable depth precision, typically 24 bits.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_WINDOW\_IATTR\_PLANES\_STENCIL\\
  \textbf{Value} & unsigned\\
  \textbf{Default} & UNDEFINED\\
  \textbf{Details} & Determines the number of stencil planes for the
  window. The interpretation of this value is window system-specific, as
  some window systems select a visual with the closest match to this
  value, and some select a visual with at least the number of stencil
  planes specified. By default no stencil planes are requested. Note
  that for depth-compositing and pixel-compositing at least one stencil
  plane is needed.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_CHANNEL\_IATTR\_HINT\_STATISTICS\\
  \textbf{Value} & OFF \textbar \ FASTEST [ ON ] \textbar \ NICEST\\
  \textbf{Default} & FASTEST [Release Build]\\
                   & NICEST [Debug Build]\\
  \textbf{Details} & See EQ\_WINDOW\_IATTR\_HINT\_STATISTICS.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_MODE\\
  \textbf{Value} & QUAD \textbar \ ANAGLYPH\\
  \textbf{Default} & QUAD\\
  \textbf{Details} & Selects the algorithm used for stereo
  rendering. QUAD-buffered stereo use the left and right buffers of a
  stereo window (active stereo). Anaglyphics stereo uses
  \textsf{glColorMask} to mask colors for individual eye passes, used in
  conjunction with colored glasses.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_LEFT\_MASK\\
  \textbf{Value} & [ RED GREEN BLUE ]\\
  \textbf{Default} & [ RED ] \\
  \textbf{Details} & Select the color mask for the left eye pass during
  anaglyphic stereo rendering.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_STEREO\_ANAGLYPH\_RIGHT\_MASK\\
  \textbf{Value} & [ RED GREEN BLUE ]\\
  \textbf{Default} & [ GREEN BLUE ]\\
  \textbf{Details} & Select the color mask for the right eye pass during
  anaglyphic stereo rendering.\\
  \hline
\end{tabularx}\\\vfill

\begin{tabularx}{\textwidth}{|l|X|}
  \hline
  \textbf{Name} & EQ\_COMPOUND\_IATTR\_UPDATE\_FOV\\
  \textbf{Value} & HORIZONTAL [ON] \textbar \ VERTICAL \textbar \ OFF\\
  \textbf{Default} & HORIZONTAL\\
  \textbf{Details} & Define how the wall or projection description is
  updated when a window is resized. Horizontal increases or decreases
  the horizontal field of view, while keeping the vertical size
  constant. Vertical updates the vertical field of view, while keeping
  the horizontal size constant. OFF does not update the wall or
  projection description at all.\\
  \hline
\end{tabularx}\\
\end{center}

@@

The server section consists of connection description parameters for the
server listening sockets, and a number of configurations for this
server. Currently only the first configuration is used.

\subsubsection{\label{sConnectionDescription}Connection Description}

A connection description defines the network parameters of an Equalizer
process. Currently two connection types, TCP/IP and SDP, are
supported. TCP/IP creates a TCP socket. SDP is very similar, except that
the address family \textsf{AF\_INET\_SDP} instead of \textsf{AF\_INET}
is used to enforce a SDP connection. Note that you can also use the
transparent mode provided by most InfiniBand implementations to used SDP
with TCP connections.

Furthermore, a port for the socket can be specified. When no port is
specified for the server, the default port 4242 (+UID on Posix systems)
is used. When no port is specified for a node, a random port will be
chosen by the operating system. For pre-launched render clients, a port
has to be specified for the server to find the client node.

The hostname is the IP address or (resolvable) host name. When used for
the server, the server binds its listening socket only to the specified
name.

A server or node may have multiple connection descriptions.
\clearpage
{\footnotesize\begin{lstlisting}
server
{
    connection // 0-n times, listening connections of the server
    {
        type       TCPIP | SDP
        TCPIP_port unsigned
        hostname   string
    }

@@

@@

A configuration has a number of parameters, nodes and compounds.

The nodes and their children describe the rendering resources in a
natural, hierarchical way. Compounds use rendering resources (channels)
to execute rendering tasks. For an introduction on the concepts of
nodes, pipes, windows, channels and compounds please refer to
\sref{sConfig} and \sref{sCompounds}.

The latency of a config defines the maximum number of frames the
slowest operation may fall behind the application thread. A latency of 0
synchronizes all rendering tasks started by \textsf{Config::startFrame}
in \textsf{Config::finishFrame}. A latency of one synchronizes all
rendering tasks started one frame ago in \textsf{finishFrame}.

For a description of config attributes please refer to \sref{sGlobal}.

@@

For a description of connection attributes please refer to \sref{sGlobal}.

{\footnotesize\begin{lstlisting}
node
{
    name     string
    connection // 0-n times, possible connections to this node
    {
        type          TCPIP | SDP
        TCPIP_port    unsigned
        hostname      string
        command       string      // render client launch command
        command_quote 'character' // launch command argument quote char
        timeout       unsigned    // timeout in milliseconds for launch
    }
    pipe {...} // 1-n times
}
\end{lstlisting}}
